\section{Benchmarks}
\label{sec:benchmarks}

From the \coq\ formal specification, we have extracted an Ocaml checker
implementation. In the following table, we have collected several benchmarks.
For each test, we
give the size of the two tree automata (initial $\A^0$ and completed $\A_\R^*$)
as number of transitions/number of states.  For each TRS $\R$ we give the number
of rules.  The 'CS' column gives the number of completion steps necessary to
complete $\A^0$ into $\A^*_{\R}$ and 'CT' gives the completion time.  The 'CKT'
column gives the time for the checker to certify the $\A^*_{\R}$ and the 'CKM'
gives the memory usage. The important thing to observe here is that, the
completion time is very long (sometimes more than 24 hours), the {\em checking}
of the corresponding automaton is always fast (a matter of seconds).

The four tests are Java programs translated into term rewriting systems using
the technique detailed in~\cite{BoichutGJL-RTA07}. All of them are completed
using \timbuk\ except the example {\tt List2.java} which has been completed using a completion
tool under development by Yohan Boichut and Emilie Balland.  This shows that the
completed automaton produced by a new and fully optimzed tool is also accepted
by our checker.  The {\tt List1.java} and {\tt List2.java} corresponds to the
same Java program but with slightly different encoding into TRS and
approximations. The {\tt Ex\_poly.java} is the example given
in~\cite{BoichutGJL-RTA07} and the {\tt Bad\_Fixp} is the same problem as {\tt
  Ex\_poly.java} except that the completed automaton $\A^*_{\R}$ has been
intentionally corrupted. Thus, this is thus not a valid fixpoint and rejected by
the checker.

\begin{center}
\begin{tabular}{l|c|c|c|c|c|c|c|c}
  Name & $\A^0$ & $\A^*_\R$ & $\R$ & CS & CT & CKT & CKM \\ \hline
  
  {\tt List1.java} & 118/82 & 422/219 & 228 & 180 & $\approx$ 3 days & 0,9s & 2,3 Mo \\ \hline
  %Yoh
  {\tt List2.java} & 1/1 & 954/364 & 308 & 473 & 1h30 & 2,2s & 3,1 Mo \\ \hline
  %Res_
  {\tt Ex\_poly.java} & 88/45 & 951/352 & 264 & 161 & $\approx$ 1 day & 2,5s & 3,3 Mo \\ \hline
  %Res_ modifié
  {\tt Bad\_Fixp} & 88/45 & 949/352 & 264 & 161 & $\approx$ 1 day & 1,6s & 3,2 Mo \\ \hline

%  NSPK & 26tr. 12 states & 243 tr. 12st. & 13 & 6 & 8'/21.4s & 5 Mo \\ \hline
\end{tabular}
\end{center}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: 
