\documentclass[a4paper,10pt]{llncs}
\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage[curve]{xypic}
\usepackage{stmaryrd}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{bussproofs}
\setlength{\oddsidemargin}{15mm}
\setlength{\evensidemargin}{15mm} 
\setlength{\textwidth}{135mm}
\setlength{\textheight}{214mm} 
\setlength{\topmargin}{5mm}

\input{rapinclude}

%All i need to have fun with Caml and Coq :)
\RequirePackage{alltt}
\RequirePackage{listings}
\def\lstlanguagefiles{lstlangcoq.sty,lstlangcaml.sty}
\lstloadlanguages{Coq,Caml} %Caml
\newcommand{\switchlstcoq}{
\lstset{language=Coq,flexiblecolumns=false,mathescape=true}
\lstset{keywordstyle={\bfseries}}
\lstset{commentstyle=\it,basicstyle=\small\tt}%,keywordstyle=\bfseries}
\lstset{literate={=>}{{$\Rightarrow$}}1
{\\Abstate}{{\Abstate}}1
{forall}{{$\forall$\hspace{-1ex}}}1
{exists}{{$\exists$\hspace{-1ex}}}1
{not}{{$\neg$}}1
{<=<}{{$\subseteq$}}2
{>=}{{$\ge$}}1
{>->}{{$\rightarrowtail$}}2{->}{{$\to$}}2 {/\\}{{$\land$}}1 {~}{{$<AC>$}}1}
\lstset{escapeinside={(*@}{@*)}} % In the text as an example:
}

\renewcommand{\ttdefault}{pcr}


\begin{document}
\versioncourte{\title{Certifying a Tree Automata Completion Checker\thanks{For
      reviewer's convenience, proofs of~\cite{BoyerGJ-RR08} are given as an
      appendix.}}}  \versionlongue{\title{Certifying a Tree Automata Completion
    Checker}} \author{Benoît Boyer \and Thomas Genet \and Thomas Jensen}

\institute{
  IRISA / Universit\'e de Rennes\,1 / CNRS \\
  Campus de Beaulieu \\
  F-35042 Rennes Cedex \\
  {\tt {bboyer,genet,jensen}@irisa.fr} 
}
\maketitle


\begin{abstract}
  Tree automata completion is a technique for the verification of infinite state
  systems. It has already been used for the verification of cryptographic
  protocols and the prototyping of Java static analyzers.  However, as for many
  other verification techniques, the correctness of the associated tool becomes
  more and more difficult to guarantee. It is due to the size of the
  implementation that constantly grows and due to optimizations which
  are necessary to scale up the efficiency of the tool to verify real-size
  systems. In this paper, we define and develop a checker for tree automata
  produced by completion. The checker is defined using \coq\ and its
  implementation is automatically extracted from its formal specification. Using
  extraction gives a checker that can be run independently of the \coq\
  environment. A specific algorithm for tree automata inclusion checking have
  been defined so as to avoid the exponential blow up. The obtained checker is
  certified in \coq, independent of the implementation of completion, usable
  with any approximation performed during completion, small and fast. Some
  benchmarks are given to show how efficient the tool is.
\end{abstract}

%each file contains a section :


\input{intro}

\input{preliminaries}

\input{completion}

\input{objectives}

\input{rewriting}

\input{automata}

\input{inclusion}

\input{closure}

\input{benchs}

%\section{Complete certificate}
%\label{sec:certificate}
%Peut-être une partie pour expliquer comment termine-t-on la preuve 
%pour obtenir le certificat.
\section{Conclusion and further research}
\label{sec:conclusion}
In this paper we have defined a \coq\ checker for tree automata completion.  The
first characteristic of the work presented here is that the checker does not
validate a specific implementation of completion but, instead, the result. As a
consequence, the checker remains valid even if the implementation of the
completion algorithm changes or is optimized. A second salient feature is that
the code of the checker is directly generated from the correctness proof of its
verified \coq\ specification through the \coq\ extraction mechanism. Third, we
have payed particular attention to the formalization of the checker in order not
to lose efficiency to obtain the certification. We have defined a specific
inclusion algorithm in order to avoid the usual exponential blow-up obtained
with the standard inclusion algorithm.  We have defined the \coq\ formal
specification so that it was possible to extract an independent OCaml
checker. Finally, we made an extensive use of efficient formal data structures
leading to more complex proof but also to faster extracted checker.  An
extension for non left-linear TRS, which are sometimes necessary for specifying
cryptographic protocols, is under development.  Since many different kind of
analysis can be expressed as reachability problems over tree automata, and since
verification of completed automata revealed to be very efficient, we aim at
using this technique in a PCC architecture where tree automata are viewed as
program certificates. At last, note that even if this checker is external to
\coq, we can use the correction proof of the checker and the \coq\ reflexivity
mechanism to lift-up the external verification into a proof in the \coq\
system. This can be necessary to perform efficient unreachability proofs on
rewriting systems in \coq\ using an external completion tool.


\bibliographystyle{alpha}
\bibliography{sabbrev,eureca,genet}

\newpage
\versioncourte{
\appendix
\section{Proofs}
%\annexe
%\newpage
\setcounter{theorem}{0}
\setcounter{lemma}{0}
\input{annexe}}
\end{document}
